Discrete Quantum Theories and Computing

### 2018-12-12 dissertation.tex
dissertation.lyx "Chapter 1 Introduction" Updating...

Since we as human cannot distinguish quantities differing only by
an arbitrarily small extent \cite{Turing_1937,Gisin2017}, quantum
mechanics and computation based on the continuum of complex numbers
may not capture our ability of predicting quantum phenomenon and the
power of quantum computation. Being agnostic about whether the reality
is ultimately discrete or continuous, we considered two types of models
to incorporate our limitation of distinguishability. By assuming the
quantum states are quantized and distinguishable, the continuum of
complex numbers is replaced by finite fields which serves as the basis
of the first type of our quantum theories and computing. We then shift
our attention from the reality to the limitation of distinguishability
itself, incorporate the finite precision measurement into our model,
and consider quantum interval-valued probability measures.

Even before quantum mechanics is discovered, human tried to characterize
the physical quantities and manipulate them to simulate and predict
unknown quantities. Discrete quantities can be easily characterized
by whole numbers~\cite{ParkerBaldridge2004}. Even if the proportion
of whole numbers are fractions, the operation among fractions can
be considered as the operation among whole numbers after we find the
common denominator, and the resulting numbers faithfully predicts
the corresponding quantities. In general, discrete computation can
be reliably repeated with exact equivalence, and any digital computer
can faithfully simulate a Turing machine limited only by the available
memory and time~\cite{Turing_1937}.

In contrast, the non-apparently discrete quantities confused many
cultures from the very beginning. Some people, including Greek mathematicians
in the school of Pythagoras, tried to apply their discrete idea into
their model, and believed even the mathematical models must be ultimately
discrete. In legend, they murdered Hippasus because he found there
is no minimum unit between a side of a square and its diagonal. Comparing
to admit that the length of a square's diagonal is not a physical
quantity or there is no square physically, working with a mathematical
model without minimum unit seems more appealing, and their corresponding
physical quantities are also assumed to be continuous. An ancient
Chinese philosopher Chuang-tzu (庄子) expressed this belief as ``If
a one-foot-long stick is cut into halves every day, the cutting will
never come to an end''\footnote{``一尺之棰，日取其半，萬世不竭。'' in Chinese \cite{Zhuangzi1999,Liu2018}.}.
Later, these continuous quantities becomes the basic component of
analog computers. After logarithm was discovered, a slide rule, also
known as a slipstick, is used to compute multiplication, division,
and more complex operations as a mechanical analog computer~\cite{wiki:SlideRule}.
Charge amplifiers is used to build integrators \cite{wiki:Integrator,wiki:ChargeAmplifier}.

@@

Prediction\tabularnewline[\doublerulesep]

@@

of Events\tabularnewline[\doublerulesep]

@@

\end{tabular} & Defined &  &  & Undefined &  & Defined\tabularnewline\addlinespace

>>>>

Since we as human cannot distinguish quantities differing only by
an arbitrarily small extent \cite{Turing_1937,Gisin2017}, quantum
mechanics and computation based on the continuum of complex numbers
may not capture our ability of predicting quantum phenomenon and the
power of quantum computation. Being agnostic about whether the reality
is ultimately discrete or continuous, we considered two types of models
to incorporate our limitation of distinguishability into quantum theories.
By assuming the quantum states are ultimately quantized and distinguishable,
the first type of our quantum theories replaces the complex numbers
by discrete finite fields. We then shift our attention to focus on
the limitation of distinguishability itself, incorporate the finite
precision measurement into our model, and consider quantum interval-valued
probability measures.

Even before quantum mechanics is discovered, human tried to characterize
the physical quantities and manipulate them to simulate and predict
unknown quantities. Discrete quantities can be easily characterized
by whole numbers~\cite{ParkerBaldridge2004}. Even if the proportion
of whole numbers are fractions, the operations among fractions is
essentially the same as those among whole numbers over the common
denominator, and the computational results faithfully represent the
corresponding quantities. In general, discrete computation can be
reliably repeated with exact equivalence, and any digital computer
can faithfully simulate a Turing machine limited only by the available
memory and time~\cite{Turing_1937}.

In contrast, the non-apparently discrete quantities confused many
cultures since very beginning. For example, Greek mathematicians in
the school of Pythagoras tried to apply their discrete idea into their
model, and believed every mathematical model must be ultimately discrete.
They believed so much that they in legend murdered Hippasus because
he found there is no minimum unit between a side of a square and its
diagonal. Comparing the idea that the length of a square's diagonal
is not a physical quantity or there is no square physically, working
with a mathematical model without minimum unit seems more appealing,
and their corresponding physical quantities are also assumed to be
continuous. An ancient Chinese philosopher Chuang-tzu (庄子) expressed
this belief as ``If a one-foot-long stick is cut into halves every
day, the cutting will never come to an end''\footnote{``一尺之棰，日取其半，萬世不竭。'' in Chinese \cite{Zhuangzi1999,Liu2018}.}.
Later, these continuous quantities becomes the basic component of
analog computers. After logarithm was discovered, a slide rule, also
known as a slipstick, is used to compute multiplication, division,
and more complex operations as a mechanical analog computer~\cite{wiki:SlideRule}.
Charge amplifiers is used to build integrators \cite{wiki:Integrator,wiki:ChargeAmplifier}.

@@

Likelihood of\tabularnewline[\doublerulesep]

@@

Events is\tabularnewline[\doublerulesep]
Predicted by\tabularnewline[\doublerulesep]

@@

\end{tabular} & Defined & Undefined & Undefined & Undefined &  & Defined\tabularnewline\addlinespace
Contextuality & Yes & Supercontextuality &  &  &  & %
Yes when Measurement\tabularnewline[\doublerulesep]
Precise Enough\tabularnewline[\doublerulesep]
